---
title: "244 Lab 6"
author: "Molly Williams"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Goals: 

- Read in spatial data (shapefiles)
- Simplify polygons (st_simplify)
- Add and transform projection (st_transform and st_crs)
- Create several finalized-ish maps in R
- Join spatial data (st_join)
- Find spatial intersections (st_intersect)
- Interactive plots with tmap
- Updating color schemes, base layers, etc.
- Plotting spatial lines
- Creating sf data from lat/lon coordinates

Awesome resource: 
*Geocomputation in R* by Robin Lovelace, available online: 
<https://geocompr.robinlovelace.net/>
```{r}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(ggrepel)
library(ggspatial)
library(RColorBrewer)
library(raster)
```

####Example 1: Dams in California

```{r}
# sf package is for reading shape files

ca_eco <- read_sf(".", layer = "ca_eco") %>%  # pulls all the shape files with the ca_eco prefix
  dplyr::select(US_L3NAME) %>% # Only select column with eco-regions
  rename(Region = US_L3NAME) %>% # Rename that column to "Region"
  st_simplify(dTolerance = 100) %>% # Simplify polygons for time-saving purposes; within 100 meters gets collapsed onto a single line
  st_transform(crs = 4326) # Change coordinate reference system (CRS) to 4326
  

```


Read in CA Counties shapefile data and dam data:
```{r}

ca_counties <- read_sf(".", layer = "california_county_shape_file")

st_crs(ca_counties) = 4326

ca_dams <- read_sf(dsn = ".", layer = "California_Jurisdictional_Dams") %>% # Read data
  rename(Condition = Condition_) # Change column name (remove final _)

ca_dams$Condition <- fct_relevel(ca_dams$Condition, "Fair","Satisfactory","Unsatisfactory","Poor") # Set factor levels (not sure if using this later...)

```



### Make maps!!

```{r}

plot(ca_eco)
plot(ca_counties)


```

### Make a map with ggplot:

```{r}

# We have more eco regions than colors in the color palette

color_count <- 13 #there are 13 eco regions 

my_colors <- colorRampPalette(brewer.pal(10, "Set2"))(color_count) # 10 = number of colors in existing palette, set2 syntax changes number to the # specified in color_count 

ggplot(ca_eco)+
  geom_sf(aes(fill = Region), 
          color = "NA",
          show.legend = FALSE) +
  scale_fill_manual(values = my_colors) +
  geom_sf(data = ca_counties, 
          fill = "NA", # just shows the county outlines
          color = "gray30",
          size = 0.1) +
  geom_point(data = ca_dams, 
             aes(x = Longitude, y = Latitude), 
             size = 1, 
             color = "gray10", 
             alpha = 0.5) +
  theme_minimal() +
  coord_sf(datum=NA)
```



### Example 2: Dams in the Sierra Nevada Eco-region

```{r}

sn <- ca_eco %>% 
  filter(Region == "Sierra Nevda") %>% 
  st_join(ca_dams)

ggplot(sn) +
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA") +
  geom_sf(fill = "lemonchiffon4", color = "NA") +
  geom_point(aes(x = Longitude, y = Latitude),
             size = 0.5,
             color = "red4")


```

### Example 3: Eco-regions for Santa Barbara County
```{r}

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara") # Results in four rows because in SBC we have islands!

# Create clipped version of eco-region data that only includes its intersection with the sb polygon 

eco_clip <- st_intersection(ca_eco, sb)
plot(eco_clip) # shows data within the bounds of SBC 

# But... looks better in ggplot, of course:

ggplot(eco_clip) +
  geom_sf(data = ca_counties, 
          fill = "gray90", 
          color = "gray80", 
          size = 0.2) + # regions just in SB county highlighted; 1st layer
  geom_sf(aes(fill = Region), color = "NA") + #shows eco regions for SB county on top of CA basemap, but let's make it more awesome
  scale_fill_manual(values = c("darkolivegreen2","darkolivegreen","gold2")) +
  coord_sf(xlim = c(-121, -119), ylim = c(33.5, 35.5)) + #crops to region of interest
  geom_point(aes(x = -119.6982, y = 34.4208), size = 2) + #adds point for SB
  geom_text (x = -119.6982, y = 34.35, label = "Santa Barbara") +
  theme_minimal() + # Update theme
  theme(legend.position = c(0.5,0.15)) +# Move the legend
  labs(x = "", y = "", title = "Santa Barbara County Eco-Regions")


```

### Example 4: Intro to Interactive Plots with tmap

```{r}

map_sb_eco <- tm_shape(eco_clip) +
  tm_fill("Region", palette = "RdPu", alpha = 0.5) +
  tm_shape(ca_counties) + 
  tm_borders()

tmap_mode("view")
map_sb_eco

# leafleft::providers

tm_basemap("CartoDB.DarkMatter")+
  tm_shape(eco_clip)+
  tm_borders(col = "white")
```


### Example 5: Fault Lines

```{r}

fault_lines <- read_sf(dsn = ".", layer = "GMC_str_arc") %>% 
  st_transform(crs = 4326) %>% 
  separate(LTYPE, into = c("syn_ant", "certainty", "direction"), sep = ",")

# Syncline or anticline (down or up); scenarios are broken up but geometry is retained 

# Base plot:
plot(fault_lines)

# All CA: 

ggplot() +
  geom_sf(data = ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  theme_dark()

# Limit to faults within SB polygon: 
sb_faults <- fault_lines %>% 
  st_intersection(sb) 

# Plotting with ggplot: 
ggplot() +
  geom_sf(data = sb) +
  geom_sf(data = sb_faults, aes(color = syn_ant))

# Plotting with tmap: 

tmap_mode("view")
tm_basemap("CartoDB.DarkMatter") +
  tm_shape(sb) +
  tm_borders(col = "gray50", lwd = 2) +
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange","purple"), lwd = 2)

```

### Example 6: Faceted maps

```{r}

ggplot() +
  geom_sf(data = ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  theme_dark() +
  facet_wrap(~syn_ant) # Choose variable to facet by


# Can also do this with tmap: 

tm_basemap("CartoDB.DarkMatter") +
  tm_shape(sb) +
  tm_borders(col = "gray50", lwd = 2) +
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange","purple"), lwd = 2) +
  tm_facets(by = "syn_ant")

```


### Example 7. Making spatial points from latitude/longitude information

```{r}

ca_sites <- read_csv("cadfw_sensitive_sites.csv")


# Read in by longitude and latitude in CSV, and set CRS

sites_sf <- st_as_sf(ca_sites, coords = c("LONDD","LATDD"), crs = 4326)

# Then make a plot: 

ggplot() +
  geom_sf(data = ca_counties, fill = "gray40") +
  geom_sf(data = sites_sf, aes(color = PRI_CODE), size = 0.5)

# Sticky geometries ! 



```


### Example 8: Chloropleth of CA counties by NUMBER of dams in each county
*Use dplyr functions with spatial information*

```{r}

intersection <- st_intersection(ca_dams, ca_counties)

# Count how many dams there are in each county:

dams_per_county <- intersection %>% 
  group_by(NAME) %>% 
  tally()

```









